---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file. -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.path  = "man/figures/README-"
)
library(shooters)
```

# shooters

`shooters` provides static-analysis utilities for Shiny applications.  The
flagship function, `ns_tree()`, parses R source files in a directory and
prints a plain-text call tree showing how functions relate to one another —
with no need to run the app.

## Installation

```r
# install.packages("pak")
pak::pak("mjfrigaard/shooters")
```

## Usage

Point `ns_tree()` at any directory that contains R source files:

```r
ns_tree("R")
```

### Example apps

`shooters` bundles four example apps that cover the most common Shiny
structural patterns.

```{r apps-path}
apps <- system.file("apps", package = "shooters")
```

#### `minimal` — bare `ui` / `server`

No wrapper functions and no modules. `ui` is a plain object; `server` is the
only defined function. Use `ui_fun` / `server_fun` to tell `ns_tree()` what
to look for:

```{r minimal}
ns_tree(file.path(apps, "minimal"), ui_fun = "ui", server_fun = "server")
```

#### `no_modules` — helper functions, no modules

A `launch()` entry point calls `app_ui()` and `app_server()`. The server
delegates to two plain helper functions:

```{r no-modules}
ns_tree(file.path(apps, "no_modules"))
```

#### `single_module` — one NS / moduleServer pair

A scatter-plot module (`scatter_ui` / `scatter_server`) is nested inside the
app-level UI and server:

```{r single-module}
ns_tree(file.path(apps, "single_module"))
```

#### `nested_modules` — modules calling other modules

A `display` parent module delegates to `plot` and `table` child modules,
spread across multiple files:

```{r nested-modules}
ns_tree(file.path(apps, "nested_modules"))
```

## How it works

`ns_tree()` operates in four steps:

1. **Parse** every `.R` file in the target directory with `parse()`.
2. **Extract** top-level function definitions (`name <- function(...)`).
3. **Walk** each function body to find references to other locally defined
   functions, building a call graph.
4. **Render** the call graph as a depth-first ASCII tree, starting from the
   app root.

The root is resolved with a three-tier fallback:

| Tier | Condition | Root used |
|------|-----------|-----------|
| 1 | `app_fun` (default `"launch"`) is defined | That function |
| 2 | `ui_fun` or `server_fun` is defined | Co-roots under `(app)` |
| 3 | None of the above | All defined functions under `(app)` |

## Demo apps

Run any of the bundled example apps interactively:

```r
run_demo()                    # nested_modules (default)
run_demo("single_module")
run_demo("no_modules")
run_demo("minimal")
```
