---
title: "Getting started with shooters"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with shooters}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>"
)
library(shooters)
```

`shooters` provides `ns_tree()`, a static-analysis tool that reads R source
files from a directory, builds a call graph of every function defined there,
and prints a plain-text tree showing how those functions call one another.
It is designed for Shiny apps — where understanding namespace structure matters
— but it works on any collection of R source files.

## How `ns_tree()` works

`ns_tree()` performs four steps on every `.R` file it finds in the target
directory:

1. **Parse** — `parse()` turns each file into an expression list.
2. **Extract** — `extract_func_def()` identifies top-level function
   assignments (`name <- function(...) { ... }`).
3. **Walk** — `find_calls()` recursively walks each function body and records
   references to other *known* (i.e. locally defined) functions.
4. **Render** — `build_tree()` traverses the resulting call graph depth-first
   from a root node, and `render_tree()` formats the nested list as
   box-drawing ASCII art.

The root node is resolved with a three-tier fallback:

| Tier | Condition | Behaviour |
|------|-----------|-----------|
| 1 | `app_fun` is defined | Use it as the single tree root |
| 2 | `ui_fun` or `server_fun` is defined | Co-roots under a synthetic `(app)` node |
| 3 | None of the above | All defined functions listed flat under `(app)` |

---

## Example apps

`shooters` bundles four example apps in `inst/apps/` that cover the most
common Shiny structural patterns. The path to each app directory can be
retrieved with `system.file()`.

```{r app-base}
apps <- system.file("apps", package = "shooters")
```

### 1. `minimal` — bare `ui` / `server`, no functions

The simplest possible Shiny app: `ui` is a plain R object (not a function),
`server` is a function defined at the top level, and both are passed directly
to `shinyApp()`. There is no launcher wrapper.

```r
# inst/apps/minimal/app.R
ui <- fluidPage(
  titlePanel("Minimal App"),
  sidebarLayout(
    sidebarPanel(sliderInput("n", "Number of points", min = 10, max = 200, value = 50)),
    mainPanel(plotOutput("plot"))
  )
)

server <- function(input, output) {
  output$plot <- renderPlot({
    d <- data.frame(x = rnorm(input$n), y = rnorm(input$n))
    plot(d$x, d$y, pch = 19, col = "steelblue")
  })
}

shinyApp(ui, server)
```

Because there is no `launch()` function, `ns_tree()` falls through to **Tier
2**: it looks for `ui_fun` and `server_fun` by name. `ui` is not a function
definition so `extract_func_def()` skips it; only `server` is captured. Pass
the actual names used in the file via the `ui_fun` / `server_fun` arguments:

```{r minimal}
ns_tree(
  file.path(apps, "minimal"),
  ui_fun     = "ui",
  server_fun = "server"
)
```

---

### 2. `no_modules` — helper functions, no Shiny modules

The app is split into a `launch()` entry point, `app_ui()`, `app_server()`,
and two plain helper functions (`make_data()`, `render_scatter()`). No
`NS()` or `moduleServer()` are used.

```r
# inst/apps/no_modules/app.R  (selected functions)
app_server <- function(input, output, session) {
  plot_data <- make_data(input)
  output$scatter <- render_scatter(plot_data)
}

make_data <- function(input) { ... }
render_scatter <- function(data) { ... }

launch <- function() {
  shinyApp(ui = app_ui(), server = app_server)
}
```

`launch()` is found immediately (**Tier 1**), so the tree is rooted there.
The helper calls inside `app_server` appear as its children:

```{r no-modules}
ns_tree(file.path(apps, "no_modules"))
```

---

### 3. `single_module` — one NS / moduleServer pair

A scatter-plot module (`scatter_ui` / `scatter_server`) is defined in the
same file and called from the app-level `app_ui()` / `app_server()`.

```r
# inst/apps/single_module/app.R  (selected functions)
scatter_ui <- function(id) {
  ns <- NS(id)
  tagList(sliderInput(ns("n"), ...), plotOutput(ns("plot")))
}

scatter_server <- function(id) {
  moduleServer(id, function(input, output, session) { ... })
}

app_ui <- function() {
  fluidPage(titlePanel("Single Module App"), scatter_ui("scatter1"))
}

app_server <- function(input, output, session) {
  scatter_server("scatter1")
}

launch <- function() shinyApp(ui = app_ui(), server = app_server)
```

The tree shows the module pair nested one level below `app_ui` / `app_server`:

```{r single-module}
ns_tree(file.path(apps, "single_module"))
```

---

### 4. `nested_modules` — modules calling other modules

The most realistic pattern: a `display` parent module owns a slider and
delegates rendering to two child modules (`plot`, `table`). Each module pair
lives in its own file; `app.R` sources them all.

```
inst/apps/nested_modules/
├── app.R          # launch(), sources everything
├── app_ui.R       # app_ui()  → display_ui()
├── app_server.R   # app_server() → display_server()
├── mod_display.R  # display_ui(), display_server() → plot_*, table_*
├── mod_plot.R     # plot_ui(), plot_server()
└── mod_table.R    # table_ui(), table_server()
```

`ns_tree()` reads *all* `.R` files in the directory so the full cross-file
call graph is assembled before the tree is built:

```{r nested-modules}
ns_tree(file.path(apps, "nested_modules"))
```

The tree faithfully mirrors the physical module hierarchy: `display` is the
parent that bridges the app level to the `plot` and `table` leaf modules.

---

## Key arguments

```r
ns_tree(
  path       = "R",          # directory to scan (default: "R/")
  app_fun    = "launch",     # Tier-1 root function name
  ui_fun     = "app_ui",     # Tier-2 UI co-root name
  server_fun = "app_server"  # Tier-2 server co-root name
)
```

Adjust `app_fun`, `ui_fun`, and `server_fun` to match whatever naming
convention the target app uses.

---

## Running the demo apps

`run_demo()` launches any of the four example apps interactively:

```r
run_demo()                    # nested_modules (default)
run_demo("single_module")
run_demo("no_modules")
run_demo("minimal")
```
